Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2012-08-08 paper/paper.tex
add ref to simulator
Add Ewa's changes

Workflows in an ensemble may differ not only in their parameters, but also in
their priority. For example, in CyberShake some sites may be in heavily
populated areas or in strategic locations such as power plants, while others
may be less important. Scientists typically prioritize the workflows in such
an ensemble so that important workflows are finished first. This enables them
to see critical results early, and helps them to choose the most important
workflows when the time and financial resources available for computing are
limited.

@@

The slack time of a level $l$ in workflow $w$ is given by:
%
$$
ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
$$
%
where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of
tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow,
$R(w)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a
parameter between 0 and 1 that causes more slack time to be given to levels
with more tasks (large $\alpha$) or more runtime (small $\alpha$).

@@

To evaluate and compare the three proposed algorithms, we developed a cloud
workflow simulator based on CloudSim~\cite{Calheiros2011}. Our simulation model
consists of Cloud, VM and WorkflowEngine entities. The Cloud entity starts and
terminates VM entities using an Amazon EC2-like API. VM entities simulate the
execution of individual tasks, including randomized variations in runtime. The
WorkflowEngine entity manages the scheduling of tasks and the provisioning of 
VMs based on the chosen algorithm. We assume that the VMs have a single core
and execute tasks sequentially. 
%Although CloudSim provides a more advanced infrastructure model which includes
%time-sharing and space-sharing policies, we do not use these features since we
%are interested mainly in the execution of tasks on VMs and high-level workflow
%scheduling and provisioning. 
The simulator reads workflow description files in
a modified version of the DAX format used by the Pegasus Workflow Management
System~\cite{Deelman2005}. 
%The modified file format was created for the
%synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.

@@

Fig.~\ref{fig:distributions} shows the percentage of simulations for which
each algorithm achieved the highest score for a given ensemble type. This
experiment was conducted using all five applications, with all five types of
ensembles. For each application and ensemble type, 10 random ensembles of 50
workflows each were created. Each ensemble was simulated with all three
algorithms using 10 budgets and 10 deadlines (1000 simulations per
application, ensemble type, and algorithm). The best scores percentage is
computed by counting the number of times that a given algorithm achieved the
highest score and dividing by 1000. Note that it is possible for multiple
algorithms to get the same high score (to tie), so the numbers do not
necessarily add up to 100\%. The sum is much higher than 100\% in cases where
the dynamic algorithms perform relatively well because DPDS and WA-DPDS, which
are very similar algorithms, often get the same high score.

@@

\begin{figure*}[tb]
    \centering
    \subfloat[Ratio of Simulated Cost to Budget]{
        \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-cost-all}
        }
    \hspace{2cm}
    \subfloat[Ratio of Simulated Makespan to Deadline]{
        \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-makespan-all}
        }
    \caption{Boxplots for budget and deadline ratios when runtime estimate 
    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. Values 
    greater than 1 indicate that the budget/deadline constraint was exceeded.}
    \label{fig:variances}
\end{figure*}

Both of the workflow-aware algorithms rely on estimates of task runtimes to make
better scheduling and provisioning decisions. Our experience suggests that such
assumption is often reasonable, since we can obtain workflow performance
characteristics from preliminary runs~\cite{Bharathi2008,Deelman2005,Juve2010}.
Some applications, like Periodograms~\cite{Vockler2011} include even a
performance model that estimates task runtimes and automatically annotates
workflow description with these data. In practice, however, these
estimates are often inaccurate. Given inaccuate estimates, the question is: How
do errors in task runtime estimates impact the performance of our scheduling and
provisioning algorithms?  To examine this we introduced uniform errors in the
task runtime and observed the behavior of the algorithms in terms of meeting the
desired budget and deadline constraints.

@@

Fig.~\ref{fig:variances}.a shows the ratio of simulated cost to budget. This
plot illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions ahead of the execution and is not able to adapt
to changing circumstances.

@@

Fig.~\ref{fig:variances}.b shows the ratio of simulated makespan to budget.
Interestingly, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of the
fact that they can adapt to poor estimates and stop submitting new tasks when
the constraints are reached. Makespan is then computed as the finish time of the
last fully completed workflow from the ensemble. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a particularly
clever optimization. The SPSS algorithm tends to schedule workflows early, using
up the budget long before the deadline is reached. This is a consequence of the
deadline distribution function in SPSS, which prevents workflows from starting
late. This is illustrated by the gantt chart in
Fig.~\ref{fig:algorithm-example}.b, which shows how SPSS tends to pile up
workflows at the beginning of the timeline.
% In comparison, the dynamic algorithms tend to spread out workflow starts over
% the entire duration as shown in Fig.~\ref{fig:algorithm-example}.a. The
% consequence of this behavior is that SPSS plans tend to use up the budget but
% leave plenty of time before the deadline.
As a result, when the runtime of the plan is increased by introducing errors,
the SPSS plan has some room to expand without exceeding the deadline.

@@

\begin{figure*}[tb]
    \centering
    \subfloat[Ratio of Simulated Ensemble Cost to Budget]{
        \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-cost-montage}
        }
    \hspace{2cm}
    \subfloat[Ratio of Simulated Ensemble Makespan to Deadline]{    
        \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
        }
    \caption{Boxplots for budget and deadline ratios when provisioning delay 
    varies from 0 seconds to 15 minutes for all three algorithms. Values 
    greater than 1 indicate that the budget/deadline constraint was exceeded.}
    \label{fig:delays}
\end{figure*}

>>>>

Workflows in an ensemble may differ not only in their parameters, but also in
their priority. For example, in CyberShake some geographic locations may be in heavily
populated areas or in strategic locations such as power plants, while others
may be less important. Scientists typically prioritize the workflows in such
an ensemble so that important workflows are finished first. This enables them
to see critical results early, and helps them to choose the most important
workflows when the time and financial resources available for computing are
limited.

@@

The slack time of a level $l$ in workflow $w$ is given by:
%
$$
ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
$$
%
where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of
tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow,
$R(l)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a
parameter between 0 and 1 that causes more slack time to be given to levels
with more tasks (large $\alpha$) or more runtime (small $\alpha$).

@@

To evaluate and compare the three proposed algorithms, we developed a cloud
workflow simulator~\cite{CloudWorkflowSimulator} based on CloudSim~\cite{Calheiros2011}. Our
simulation model consists of Cloud, VM and WorkflowEngine entities. The Cloud entity starts and
terminates VM entities using an Amazon EC2-like API. VM entities simulate the
execution of individual tasks, including randomized variations in runtime. The
WorkflowEngine entity manages the scheduling of tasks and the provisioning of 
VMs based on the chosen algorithm. We assume that the VMs have a single core
and execute tasks sequentially. 
%Although CloudSim provides a more advanced infrastructure model which includes
%time-sharing and space-sharing policies, we do not use these features since we
%are interested mainly in the execution of tasks on VMs and high-level workflow
%scheduling and provisioning. 
The simulator reads workflow description files in
a modified version of the DAX format used by the Pegasus Workflow Management
System~\cite{Deelman2005}. 
%The modified file format was created for the
%synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.

@@

Fig.~\ref{fig:distributions} shows the percentage of simulations for which
each algorithm achieved the highest score for a given ensemble type. This
experiment was conducted using all five applications, with all five types of
ensembles. For each application and ensemble type, 10 random ensembles of 50
workflows each were created. Each ensemble was simulated with all three
algorithms using 10 budgets and 10 deadlines (1,000 simulations per
application, ensemble type, and algorithm, or 75,000 in total). The best scores percentage is
computed by counting the number of times that a given algorithm achieved the
highest score and dividing by 1000. Note that it is possible for multiple
algorithms to get the same high score (to tie), so the numbers do not
necessarily add up to 100\%. The sum is much higher than 100\% in cases where
the dynamic algorithms perform relatively well because DPDS and WA-DPDS, which
are very similar algorithms, often get the same high score.

@@

\begin{figure*}[tb]
    \centering
    \subfloat[Ratio of Simulated Cost to Budget]{
        \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-cost-all}
        \label{fig:variances-budget}
    }
    \hspace{2cm}
    \subfloat[Ratio of Simulated Makespan to Deadline]{
        \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-makespan-all}
        \label{fig:variances-deadline}
    }
    \caption{Boxplots for budget and deadline ratios when runtime estimate 
    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. Values 
    greater than 1 indicate that the budget/deadline constraint was exceeded.}
    \label{fig:variances}
\end{figure*}

Both of the workflow-aware algorithms rely on estimates of task runtimes to make
better scheduling and provisioning decisions. Our experience suggests that such
assumption is often reasonable, since we can obtain workflow performance
characteristics from preliminary runs~\cite{Bharathi2008,Deelman2005,Juve2010}.
Some applications, like Periodograms~\cite{Vockler2011} even include a
performance model that estimates task runtimes and automatically annotates
workflow description with these data. In practice, however, these
estimates are often inaccurate. Given inaccuate estimates, the question is: How
do errors in task runtime estimates impact the performance of our scheduling and
provisioning algorithms?  To examine this we introduced uniform errors in the
task runtime and observed the behavior of the algorithms in terms of meeting the
desired budget and deadline constraints.

@@

Fig.~\ref{fig:variances-budget} shows the ratio of simulated cost to budget. This
plot illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions ahead of the execution and is not able to adapt
to changing circumstances.

@@

Fig.~\ref{fig:variances-deadline} shows the ratio of simulated makespan to budget.
Interestingly, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of the
fact that they can adapt to poor estimates and stop submitting new tasks when
the constraints are reached. Makespan is then computed as the finish time of the
last fully completed workflow from the ensemble. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a particularly
clever optimization. The SPSS algorithm tends to schedule workflows early, using
up the budget long before the deadline is reached. This is a consequence of the
deadline distribution function in SPSS, which prevents workflows from starting
late. This is illustrated by the gantt chart in
Fig.~\ref{fig:algorithm-example-spss}, which shows how SPSS tends to pile up
workflows at the beginning of the timeline.
% In comparison, the dynamic algorithms tend to spread out workflow starts over
% the entire duration as shown in Fig.~\ref{fig:algorithm-example-dpds}. The
% consequence of this behavior is that SPSS plans tend to use up the budget but
% leave plenty of time before the deadline.
As a result, when the runtime of the plan is increased by introducing errors,
the SPSS plan has some room to expand without exceeding the deadline.

@@

\begin{figure*}[tb]
    \centering
    \subfloat[Ratio of Simulated Ensemble Cost to Budget]{
        \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-cost-montage}
        \label{fig:delays-budget}
    }
    \hspace{2cm}
    \subfloat[Ratio of Simulated Ensemble Makespan to Deadline]{    
        \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
        \label{fig:delays-deadline}
    }
    \caption{Boxplots for budget and deadline ratios when provisioning delay 
    varies from 0 seconds to 15 minutes for all three algorithms. Values 
    greater than 1 indicate that the budget/deadline constraint was exceeded.}
    \label{fig:delays}
\end{figure*}
